Rhizaria in the oligotrophic ocean exhibit clear temporal and vertical variability.

### 2024-06-26 manuscript/main.tex
edit manuscript

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=1in,textwidth=6.5in,right=1in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

% Add any tex header commands here
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Rhizaria in the oligotrophic ocean exhibit clear temporal and vertical variability.},
  pdfauthor={Alex Barth*; Leocadio Blanco-Berical; Rod Johnson; Joshua Stone},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\end{minipage}  \end{titlepage}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable]}{\end{tcolorbox}}\fi

\setstretch{2}
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Rhizaria are an extremely diverse super-group of single-celled organisms
consisting of several phyla including Retaria (foraminifera and
radiolaria) and Cercozoa. These organisms exist in a wide range of
habitats and are widely represented in plankton communities throughout
the global ocean. While the taxonomy of these organisms has recently
undergone several reclassifications (Biard, 2022a), their presence in
ocean ecosystems has been long known to oceanographers. Some of the
earliest records of their existence are from oceanographic expeditions
in the 19th century (Haekel, 1887). Rhizaria are unique members of the
plankton and protist community because they can reach large sizes (up to
several mm in diameter) and they construct intricate mineral skeletons
out of either silica, strontium, or calcium carbonate (Biard, 2022a;
Kimoto, 2015; Nakamura and Suzuki, 2015; Suzuki and Not, 2015). Despite
their noticeable morphology and global distribution, Rhizaria were
largely understudied throughout the 20th century. The bulk of modern
plankton research has focused on hard-bodied crustacea which are
numerically dominant and easily sampled with nets and preservatives.
Fragile organisms like Rhizaria were difficult to adequately study as
they can be destroyed through standard zooplankton sampling techniques
and preserve poorly. A number of studies in the late 1900s did employ
alternative techniques to quantify Rhizaria including diaphragm pumps
(Michaels, 1988) or blue-water SCUBA collections (Bijma et al., 1990;
Caron et al., 1995; Caron and Be, 1984). However, the bulk of Rhizaria
research was constrained to sediment traps or paleontological studies of
sediment (Boltovskoy et al., 1993; Takahashi et al., 1983). Only
recently has the advent of molecular techniques and in-situ imaging
tools ignited a renewed focus on Rhizaria in pelagic ecosystems (Caron,
2016).

The wave of new data on Rhizaria has facilitated an improved
understanding of the significance in ocean ecosystem functions. Firstly,
taxonomists have been able to greatly refine the understanding of
evolutionary relationships amongst these diverse protists (Aurahs et
al., 2009; Biard et al., 2015; Cavalier-Smith et al., 2018; Decelle et
al., 2013, 2012; rev by Biard, 2022a). DNA metabarcoding studies have
revealed insights into the distributional patterns (Biard et al., 2017;
Blanco-Bercial et al., 2022; Decelle et al., 2013; Llopis Monferrer et
al., 2022; Mars Brisbin et al., 2020; Sogawa et al., 2022), ecological
relationships (Decelle et al., 2012; Nakamura et al., 2023), and
contribution to biogeochemical fluxes (Guidi et al., 2016;
Gutierrez-Rodriguez et al., 2019). Transcriptomic and proteomic
approaches also have been used to quantify Rhizaria contribution to
community metabolism (Cohen et al., 2023). Yet, despite the excellent
taxonomic resolution provided by molecular approaches, they do not
provide a truly quantitative metric for estimating Rhizaria abundance or
biomass. In-situ imaging tools however, offer the ability to observe
organisms in the natural state and quantify their abundance (Barth and
Stone, In press; Ohman, 2019). Biard et al. (2016) utilized in-situ
imaging at a global scale to suggest Rhizaria were substantial
contributors to the ocean carbon standing stock. While more recent
calculations suggest lower carbon contribution (Laget et al., 2024),
Rhizaria nonetheless have substantial influences on biogeochemical
cycling. Due to their large sizes, ability to concentrate smaller
particles and the unique structure of their mineral skeletons, Rhizaria
have the potential to massively influence ocean biogeochemical cycling.
A number of studies have made large advances in estimating the
contribution of Rhizaria to ocean cycling of carbon (Gutierrez-Rodriguez
et al., 2019; Ikenoue et al., 2019; Lampitt et al., 2009; Stukel et al.,
2018), silica (Biard et al., 2018; Llopis Monferrer et al., 2021), and
strontium (Decelle et al., 2013). Still, Rhizaria ecological roles are
not well understood (Biard, 2022a). This is a major challenge as it is
critical to understand the ecological role of plankton to fully
incorporate them into biological oceanographic models.

The ecological role of Rhizaria in plankton communities is complicated
due to the fact different taxa can exhibit every different trophic
modes. As zooplankton, rhizaria are predominately heterotrophic (Biard,
2022a), yet their feeding modes can be quite varied. Phaeodarias (family
Cercozoa) are largely thought to be flux-feeders, collecting and feeding
on sinking particles (Nakamura and Suzuki, 2015; Stukel et al., 2019).
Alternatively, Retaria can be either exclusively heterotrophic or
mixotrophic, utilizing photosynthetic algal symbionts (Anderson, 2014;
Decelle et al., 2015). Mixotrophic foraminifera host a variety of
endosymbiont partners (Decelle et al., 2015; Lee, 2006), which are
thought to support early and adult life stages and significantly
contribute to total primary productivity (Kimoto, 2015). Still,
foraminifera are omnivorous, possibly even predominately carnivorous,
with several studies suggesting that they can be effective predators
(Anderson and Bé, 1976; Gaskell et al., 2019), mainly consuming live
copepods (Caron and Be, 1984). Radiolaria have several lineages, all of
which have some taxa well known to host symbionts (Biard, 2022b).
Amongst Radiolarias, arguably the most widespread are Collodaria who can
be either large solitary cells or form massive colonies, up to several
meters in length (Swanberg and Anderson, 1981). All known Collodaria
species host dinoflagellate symbionts (Biard, 2022b) and can contribute
substantially to primary productivity, particularly in oligotrophic
ocean regions (Caron et al., 1995; Dennett et al., 2002). This
Collodaria-symbiont association has been suggested as a reason for their
high abundances throughout the photic zone of oligotrophic environments
globally (Biard et al., 2017, 2016). A few Acantharea (Radiolaria order)
clades host algal symbionts (Biard, 2022b; Decelle et al., 2012),
notably with two clades forming an exclusive relationship with
\emph{Phaeocystis}. However, globally, Acantharea are less abundant than
Collodaria (Biard, 2022a) and contribute less to total primary
productivity (Michaels et al., 1995). This may be due to the fact
several clades of Acantharea are cyst-forming and strictly heterotrophs
(Biard, 2022b; Decelle et al., 2013). Furthermore, Mars Brisbin et al.
(2020) documented apparent predation behavior in Acantharea near the
surface, suggesting that there may be a larger reliance on carnivory.

Given the high abundances, yet diverse trophic strategies found among
Rhizaria taxa, it is reasonable to expect some form of niche
partitioning. A number of studies do suggest evidence for vertical
zonation between Rhizaria groups according to various trophic
strategies. Taxa-specific studies of Radiolarias suggest they may be
restricted to the euphotic zone (Boltovskoy, 2017; Michaels, 1988).
Although some studies report Acantharea in deeper waters (Decelle et
al., 2013; Gutiérrez-Rodríguez et al., 2022), Phaeodarias,
alternatively, are generally found in the mesopelagic where
photosynthesis cannot occur, but they can feed on sinking particles
(Stukel et al., 2018). In an imaging-based study of the whole Rhizaria
community, Biard and Ohman (2020) noted clear patterns in vertical
zonation which largely corresponded to different trophic roles. In the
oligotrophic ocean, Blanco-Bercial et al. (2022) also noted that the
protist community, including Rhizaria, partition along an autotroph and
mixotroph to heterotroph gradient with increasing depth in the water
column. Yet, few studies have made direct attempts to relate rhizaria
abundances to environmental factors (Biard and Ohman, 2020). In part,
this is due to the fact few studies have been able to sample Rhizaria in
the same location over a consistent timeframe (Boltovskoy et al., 1993;
Gutiérrez-Rodríguez et al., 2022; Hull et al., 2011; Michaels et al.,
1995; Michaels, 1988). Furthermore, no studies have utilized imaging,
arguably the best method for quantifying rhizaria, consistently
throughout the full mesopelagic. Given this lack of information, there
are many unknowns with respect to Rhizaria ecology, seasonality and
phenology across different groups.

In this study, we present a comprehensive assessment of large Rhizaria
measured for over a year from regularly occurring cruises at monthly
intervals. We utilized an in-situ imaging approach to facilitate
abundance calculations. With this dataset, we address two critical aims.
1) Quantification of large Rhizaria throughout the epipelagic (0-200m)
and mesopelagic (200-1000m) over the course of an annual cycle. These
data were collected in the Sargasso Sea, and represents the first study
of its kind in an oligotrophic system; and 2) We aim to test the
hypothesis that Rhizaria exhibit niche partitioning according to trophic
roles. This hypothesis makes several predictions, including vertical
zonation, as seen in prior studies, but also that environmental
variables related to trophic strategy will explain abundance patterns.
Specifically, autotrophic/mixotrophic taxa will correspond to variables
related to autotrophy (chl-a concentration, primary productivity, local
DO maxima) and other rhizaria will correspond to factors which promote
heterotrophy (particle concentration, flux, and local DO minima).

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{oceanographic-sampling}{%
\subsection{Oceanographic Sampling}\label{oceanographic-sampling}}

Data were collected in collaboration with the Bermuda Atlantic
Time-series Study (Lomas et al., 2013; Michaels and Knap, 1996) on board
the R/V Atlantic Explorer. Cruises were conducted at approximately
monthly intervals. Rhizaria individuals were sampled using the
Underwater Vision Profiler 5 {[}UVP5; Picheral et al. (2010){]}, a tool
which is well established to accurately quantify large Rhizaria (Barth
and Stone, 2022; Biard et al., 2016; Biard and Ohman, 2020; Drago et
al., 2022; Llopis Monferrer et al., 2022; Panaïotis et al., 2023; Stukel
et al., 2019; Stukel et al., 2018). The UVP5 was mounted to the sampling
rosette and collected data autonomously on routine casts, from which
only the downcast data are utilized. The UVP5 was deployed from
June-September 2019 then from October 2020 - January 2022, during which
time the BATS region was sampled for 3-5 days at monthly intervals.
Casts were filtered to only include data collected in the BATS region,
far offshore of Bermuda in the Sargasso Sea (approximately
31.0\(^oN\)-32.5\(^oN\), 64.25\(^oW\)-63\(^oW\); Supplemental Figure 1).
In general, casts extended to either 200m, 500m, or 1200m deep, with a
few extended into the bathypelagic (4500m). However, Rhizaria were only
typically found in large abundances throughout the epipelagic and
mesopelagic zones. As such, we limit this study to results from the
upper 1000m of the water column.

>>>>

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=1in,textwidth=6.5in,right=1in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

% Add any tex header commands here
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Rhizaria in the oligotrophic ocean exhibit clear temporal and vertical variability.},
  pdfauthor={Alex Barth*; Leocadio Blanco-Berical; Rod Johnson; Joshua Stone},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\end{minipage}  \end{titlepage}

\setstretch{2}
\section{Introduction}\label{introduction}

Rhizaria are an extremely diverse super-group of single-celled organisms
which occupy a wide range of habitats. Planktonic Rhizaria include
Foraminifera, Radiolaria (including Acantharea and Collodaria) and
Phaeodaria, all of which are observed throughout the global ocean (Biard
et al., 2016; Laget et al., 2024) . While the taxonomy of these
organisms has recently undergone several reclassifications (Biard,
2022a), their presence in ocean ecosystems has been long known to
oceanographers. Some of the earliest records of their existence are from
oceanographic expeditions in the 19th century (Haekel, 1887). Rhizaria
are unique members of the plankton and protist community because they
can reach large sizes (up to several mm in diameter) and they construct
intricate mineral skeletons out of either silica, strontium, or calcium
carbonate (Biard, 2022a; Kimoto, 2015; Nakamura and Suzuki, 2015; Suzuki
and Not, 2015). Despite their noticeable morphology and global
distribution, Rhizaria were largely understudied throughout the 20th
century. Fragile organisms like Rhizaria are difficult to adequately
study because they can be destroyed through standard zooplankton
sampling techniques and preserve poorly. A number of studies in the late
1900s did employ alternative techniques to quantify Rhizaria including
diaphragm pumps (Michaels, 1988) or blue-water SCUBA collections (Bijma
et al., 1990; Caron et al., 1995; Caron and Be, 1984). However, the bulk
of Rhizaria research was constrained to sediment traps or
paleontological studies of sediment (Boltovskoy et al., 1993; Takahashi
et al., 1983).

Recently, the advent of molecular techniques and \emph{in situ} imaging
tools have ignited a renewed focus on Rhizaria in pelagic ecosystems
(Caron, 2016). Molecular tools have shed light on Rhizaria taxonomy
(Biard, 2022a), distribution (Blanco-Bercial, 2020; Sogawa et al.,
2022), ecology (Decelle et al., 2012; Nakamura et al., 2023),
biogeochemical roles (Gutierrez-Rodriguez et al., 2019; Laget et al.,
2024), and metabolism (Cohen et al., 2023). Yet, despite the excellent
taxonomic resolution provided by molecular approaches, they do not
provide a truly quantitative metric for estimating Rhizaria abundance or
biomass. \emph{In situ} imaging tools however, offer the ability to
observe organisms in the natural state and quantify their abundance
(Barth and Stone, 2024; Ohman, 2019). Biard et al. (2016) utilized
\emph{in situ} imaging at a global scale to suggest large
(\textgreater500\(\mu\)m diameter) Rhizaria were substantial
contributors to the ocean carbon standing stock. While more recent
calculations suggest lower carbon contribution (Laget et al., 2024),
Rhizaria nonetheless have substantial influences on biogeochemical
cycling. This influence is due to their unique mineral skeletons and
ability to repackage small particles into larger, fast sinking ones
(Ikenoue et al., 2019). Thus, some Rhizaria may play substantial roles
in ocean export through their interception of sinking particles (Laget
et al., 2024; Stukel et al., 2018). However, Rhizaria ecology is poorly
understood (Biard, 2022a) and their ecological roles are likely more
than diverse than simple particle interception.

The ecological role of Rhizaria in plankton communities is complicated
due to the fact different taxa can exhibit very different trophic modes.
Many Rhizaria are strictly heterotrophic (Biard, 2022a), yet their
feeding modes can be quite varied. Phaeodaria are largely thought to be
flux-feeders, collecting and feeding on sinking particles (Gowing,
1989). Alternatively, Retaria can be either exclusively heterotrophic or
mixotrophic, utilizing photosynthetic algal symbionts (Anderson, 2014;
Decelle et al., 2015). Mixotrophic Foraminifera host a variety of
endosymbiont partners (Decelle et al., 2015; Lee, 2006), which are
thought to support early and adult life stages and significantly
contribute to total primary productivity (Kimoto, 2015). Still,
Foraminifera are omnivorous, possibly even predominately carnivorous,
with several studies suggesting that they can be effective predators
(Anderson and Bé, 1976; Gaskell et al., 2019), mainly consuming live
copepods (Caron and Be, 1984). Radiolaria have several lineages, many of
which host symbionts (Biard, 2022b). Amongst Radiolaria, arguably the
most widespread are Collodaria who can be either large solitary cells or
form massive colonies, up to several meters in length (Swanberg and
Anderson, 1981). All known Collodaria species host dinoflagellate
symbionts (Biard, 2022b) and can contribute substantially to primary
productivity, particularly in oligotrophic ocean regions (Caron et al.,
1995; Dennett et al., 2002). This Collodaria-symbiont association has
been suggested as a reason for their high abundances throughout the
photic zone of oligotrophic environments globally (Biard et al., 2017,
2016). A few Acantharea (Radiolaria order) clades host algal symbionts
(Biard, 2022b; Decelle et al., 2012), notably with two clades forming an
exclusive relationship with \emph{Phaeocystis}. However, globally,
Acantharea are less abundant than Collodaria (Biard, 2022a) and
contribute less to total primary productivity (Michaels et al., 1995).
This may be due to the fact several clades of Acantharea are
cyst-forming and strictly heterotrophs (Biard, 2022b; Decelle et al.,
2013). Furthermore, Mars Brisbin et al. (2020) documented apparent
predation behavior in Acantharea near the surface, suggesting that there
may be a larger reliance on carnivory.

Given the high abundances, yet diverse trophic strategies found among
Rhizaria taxa, it is reasonable to expect some form of niche
partitioning. A number of studies do suggest evidence for vertical
zonation between Rhizaria groups according to various trophic
strategies. Taxon-specific studies of Radiolaria suggest they may be
restricted to the euphotic zone (Boltovskoy, 2017; Michaels, 1988).
Although some studies report Acantharea in deeper waters (Decelle et
al., 2013; Gutiérrez-Rodríguez et al., 2022), Phaeodaria, alternatively,
are generally found beyond depths where light can sustain
photosynthesis, but they can feed on sinking particles (Laget et al.,
2024; Stukel et al., 2018). In an imaging-based study of the whole
Rhizaria community, Biard and Ohman (2020) noted clear patterns in
vertical zonation which largely corresponded to different trophic roles.
Protists communities, including Rhizaria, have been described to
partition along autotroph-mixotroph-heterotroph gradients with
increasing depth in the water column (Blanco-Bercial et al., 2022; Laget
et al., 2024). Yet, few studies have made direct attempts to relate
Rhizaria abundances to environmental factors (Biard and Ohman, 2020). In
part, this is due to the fact few studies have been able to sample
Rhizaria in the same location over a consistent timeframe (Boltovskoy et
al., 1993; Gutiérrez-Rodríguez et al., 2022; Hull et al., 2011; Michaels
et al., 1995; Michaels, 1988). Furthermore, very few studies have
utilized \emph{in situ} imaging, arguably the best method for
quantifying Rhizaria, to consistently throughout the full mesopelagic
(Laget et al., 2024) . Given this lack of information, there are many
unknowns with respect to Rhizaria ecology, seasonality and phenology
across different groups.

In this study, we present a comprehensive assessment of large Rhizaria
abundance measured for over a year from regularly occurring cruises at
monthly intervals. We utilized an \emph{in situ} imaging approach to
facilitate abundance calculations. With this dataset, we address two
critical aims. 1) Quantifying of large Rhizaria throughout the
epipelagic (0-200m) and mesopelagic (200-1000m) over the course of an
annual cycle. These data were collected in the Sargasso Sea, and
represents the first study of its kind in an oligotrophic system; and 2)
We aim to test the hypothesis that Rhizaria exhibit niche partitioning
according to trophic roles. This hypothesis has several implications,
including vertical zonation, as seen in prior studies, but also that
environmental variables related to trophic strategy will explain
abundance patterns. Specifically, autotrophic/mixotrophic taxa will
correspond to variables related to autotrophy (chl-a concentration,
primary productivity, local DO maxima) and other rhizaria will
correspond to factors which promote heterotrophy (particle
concentration, flux, and local DO minima).

\section{Methods}\label{methods}

\subsection{Oceanographic sampling}\label{oceanographic-sampling}

Data were collected in collaboration with the Bermuda Atlantic
Time-series Study (Lomas et al., 2013; Michaels and Knap, 1996) on board
the R/V Atlantic Explorer. Cruises were conducted at approximately
monthly intervals. Rhizaria individuals were sampled using the
Underwater Vision Profiler 5 (UVP) (Picheral et al., 2010). The UVP5 is
an in-situ camera used to capture plankton and particle images and has
been well established to accurately quantify large Rhizaria abundance
(Barth and Stone, 2022; Biard et al., 2016; Biard and Ohman, 2020; Drago
et al., 2022; Llopis Monferrer et al., 2022; Panaïotis et al., 2023;
Stukel et al., 2019; Stukel et al., 2018). The UVP5 was mounted to the
sampling rosette and collected data autonomously on routine casts, from
which only the downcast data are utilized. The UVP5 was deployed from
June-September 2019 then from October 2020 - January 2022, during which
time the BATS region was sampled for 3-5 days at monthly intervals (see
sampling details in Supplemental Table 1). Casts were filtered to only
include data collected in the BATS region, far offshore of Bermuda in
the Sargasso Sea (approximately 31.0\(^oN\)-32.5\(^oN\),
64.25\(^oW\)-63\(^oW\); Supplemental Figure 1). In general, casts
extended to either 200m, 500m, or 1200m deep, with a few extended into
the bathypelagic (4500m). However, Rhizaria were only typically found in
large abundances throughout the epipelagic and mesopelagic zones. As
such, we limit this study to results from the upper 1000m of the water
column.

A variety of biotic and abiotic data were collected during each BATS
cruise. The UVP5 provided particle count data at a high-frequency from
each cast. Particle concentration, as a proxy for prey field (Whitmore
and Ohman, 2021), was calculated from this data for all particles
184\(\mu m\) - 900\(\mu m\). The lower size range was set by what could
be reliably sampled by the UVP5's pixel resolution (\textgreater2px;
0.092mm per pixel) and the upper size range was set to below the size of
observable Rhizaria yet large enough to be inclusive of particles on
which some Rhizaria may feed. For each UVP cast supporting continuous
profiles of the CTD parameters salinity, temperature, and auxiliary CTD
channels; Dissolved Oxygen (DO), \emph{in situ} chlorophyll fluorescence
were measured at 24Hz using the BATS CTD package. On select casts,
Niskin bottles were used to collect bacterial abundance estimates (via
epifluorescence microscopy) as well as measure inorganic nutrients
(\(NO_3\), and \(Si\) as silicate/sicilic acid) at discrete depths. On
each cruise, flux estimates of total mass, organic carbon, and nitrogen
were also collected using sediment traps; in the present study we
utilized flux to the mesopelagic as the flux at 200m. Also primary
productivity was estimated through measuring \(^{14}C\) uptake rates
from \emph{in situ} incubations. For full descriptions of the BATS
sampling program and methods, see Knap et al. (1997) and Lomas et al.
(2013) for a review. Additionally, data can be viewed online
\href{https://bats.bios.asu.edu/bats-data/}{(https://bats.bios.asu.edu/bats-data/)}.

Environmental data were processed to match the format of the Rhizaria
abundance estimates (see below). CTD data were collected at higher
frequency than the UVP (24Hz vs 15Hz respectively), so these data were
averaged within matching bins to the UVP5 data. Data from Niskin bottles
were first linearly interpolated in depth at 1m resolution then time
averaged over the cruise, then subsequently averaged into matching bins
as the UVP5 data. Primary productivity estimates were depth integrated
throughout the euphotic zone (0-140m).
